---
title: "Dove Growth Rates"
author: "Kristen Orr"
date: "11/5/2021"
output: html_document
---

```{r}
rm(list=ls())
setwd("~/GitHub/WRC_Soil_Bacteria/analyses")
library("ggplot2")
```

```{r}
chickfull <- read.csv("../data/Example_FullChickMeasurements.csv",na = c(""))

ggplot(chickfull, aes(x = day, y = weight, color=treat)) +
  geom_point()+geom_smooth()
```
Write out some growth functions.  In all these functions
A = the asymptotic body mass (or size)
k = growth rate constant
t = age or time
T_i = age or time at the inﬂection

```{r}
logistic<-function(A,k,t=t,T_i){
  A/1+exp(-k*(t-T_i))
}
gompertz<-function(A,k,t,T_i){
  A*exp(-exp(-k*(t-T_i)))
}
EVF<-function(A,k,t,T_i){
  A*(1-exp(-exp(k*(t-T_i))))
}
Bertalanffy<-function(A,k,t,T_i){
  A*(1-exp(-k*(t+T_i)))^3
}
```
First I will use nonlinear least squares to get good starting values for the mle fits
```{r}
d=data.frame(y=chickfull$weight,t=chickfull$day+1,treat=chickfull$treat)
d=na.omit(d)

library(nlme)
logistic.=formula(y~ A/1+exp(-k*(t-T_i))|treat)#cant fit
Bertalanf=formula(y~ A*(1-exp(-k*(t+T_i)))^3|treat)#cant fit

gompz=formula(y~A*exp(-exp(-k*(t-T_i)))|treat)
m_g <- nlsList(gompz,start=list(A=150,k=0.5,T_i=25),data=d)
m_g

EVF=formula(y~A*(1-exp(-exp(k*(t-T_i))))|treat) 
m_E <- nlsList(EVF,start=list(A=150,k=0.5,T_i=25),data=d)
m_E
```

Now I will write MLE functions to estimate the parameters for 4 models and compare them with AIC.  I was only to successfully fit the Gompertz and the EVF models (Williams 1995).  For each model I fit the parameters to the pooled data (null model) and allowing different parameters for each treatment.  
```{r}
library(bbmle)
library(optimx)
startlist1=list(A=coef(m_g)[1,1],k=coef(m_g)[1,2],T_i=coef(m_g)[1,3],stdv=1)

mod_gomp_1=suppressWarnings(mle2(y ~ dnorm(gompertz(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist1,  
        parameters=list(A~1,k~1,T_i~1),method="bobyqa",
      skip.hessian=FALSE))

startlist2=list(A=coef(m_g)[,1],k=coef(m_g)[,2],T_i=coef(m_g)[,3],stdv=1)
mod_gomp_2=suppressWarnings(mle2(y ~ dnorm(gompertz(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist2,  
        parameters=list(A~treat,k~treat,T_i~treat),method="bobyqa",
      skip.hessian=FALSE))

startlist2a=list(A=coef(m_g)[,1],k=coef(m_g)[,2],T_i=coef(m_g)[1,3],stdv=1)
mod_gomp_2a=suppressWarnings(mle2(y ~ dnorm(gompertz(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist2a,  
        parameters=list(A~treat,k~treat,T_i~1),method="bobyqa",
      skip.hessian=FALSE))

startlist2b=list(A=coef(m_g)[,1],k=coef(m_g)[1,2],T_i=coef(m_g)[,3],stdv=1)
mod_gomp_2b=suppressWarnings(mle2(y ~ dnorm(gompertz(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist2b,  
        parameters=list(A~treat,k~1,T_i~treat),method="bobyqa",
      skip.hessian=FALSE))

startlist2c=list(A=coef(m_g)[1,1],k=coef(m_g)[,2],T_i=coef(m_g)[,3],stdv=1)
mod_gomp_2c=suppressWarnings(mle2(y ~ dnorm(gompertz(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist2c,  
        parameters=list(A~1,k~treat,T_i~treat),method="bobyqa",
      skip.hessian=FALSE))

startlist2d=list(A=coef(m_g)[,1],k=coef(m_g)[1,2],T_i=coef(m_g)[1,3],stdv=1)
mod_gomp_2d=suppressWarnings(mle2(y ~ dnorm(gompertz(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist2d,  
        parameters=list(A~treat,k~1,T_i~1),method="bobyqa",
      skip.hessian=FALSE))

startlist2e=list(A=coef(m_g)[1,1],k=coef(m_g)[,2],T_i=coef(m_g)[1,3],stdv=1)
mod_gomp_2e=suppressWarnings(mle2(y ~ dnorm(gompertz(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist2e,  
        parameters=list(A~1,k~treat,T_i~1),method="bobyqa",
      skip.hessian=FALSE))

startlist2f=list(A=coef(m_g)[1,1],k=coef(m_g)[1,2],T_i=coef(m_g)[,3],stdv=1)
mod_gomp_2f=suppressWarnings(mle2(y ~ dnorm(gompertz(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist2f,  
        parameters=list(A~1,k~1,T_i~treat),method="bobyqa",
      skip.hessian=FALSE))

startlist3=list(A=coef(m_E)[1,1],k=coef(m_E)[1,2],T_i=coef(m_E)[1,3],stdv=15)

mod_EVF_1=suppressWarnings(mle2(y ~ dnorm(EVF(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist3,  
        parameters=list(A~1,k~1,T_i~1),method="bobyqa",
      skip.hessian=FALSE))

startlist4=list(A=coef(m_E)[,1],k=coef(m_E)[,2],T_i=coef(m_E)[,3],stdv=1)
mod_EVF_2=suppressWarnings(mle2(y ~ dnorm(EVF(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist4,  
        parameters=list(A~treat,k~treat,T_i~treat),method="bobyqa",
      skip.hessian=FALSE))

startlist4a=list(A=coef(m_E)[,1],k=coef(m_E)[,2],T_i=coef(m_E)[1,3],stdv=1)
mod_EVF_2a=suppressWarnings(mle2(y ~ dnorm(EVF(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist4a,  
        parameters=list(A~treat,k~treat,T_i~1),method="bobyqa",
      skip.hessian=FALSE))

startlist4b=list(A=coef(m_E)[,1],k=coef(m_E)[1,2],T_i=coef(m_E)[,3],stdv=1)
mod_EVF_2b=suppressWarnings(mle2(y ~ dnorm(EVF(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist4b,  
        parameters=list(A~treat,k~1,T_i~treat),method="bobyqa",
      skip.hessian=FALSE))

startlist4c=list(A=coef(m_E)[1,1],k=coef(m_E)[,2],T_i=coef(m_E)[,3],stdv=1)
mod_EVF_2c=suppressWarnings(mle2(y ~ dnorm(EVF(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist4c,  
        parameters=list(A~1,k~treat,T_i~treat),method="bobyqa",
      skip.hessian=FALSE))

startlist4d=list(A=coef(m_E)[,1],k=coef(m_E)[1,2],T_i=coef(m_E)[1,3],stdv=1)
mod_EVF_2d=suppressWarnings(mle2(y ~ dnorm(EVF(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist4d,  
        parameters=list(A~treat,k~1,T_i~1),method="bobyqa",
      skip.hessian=FALSE))

startlist4e=list(A=coef(m_E)[1,1],k=coef(m_E)[,2],T_i=coef(m_E)[1,3],stdv=1)
mod_EVF_2e=suppressWarnings(mle2(y ~ dnorm(EVF(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist4e,  
        parameters=list(A~1,k~treat,T_i~1),method="bobyqa",
      skip.hessian=FALSE))

startlist4f=list(A=coef(m_E)[1,1],k=coef(m_E)[1,2],T_i=coef(m_E)[,3],stdv=1)
mod_EVF_2f=suppressWarnings(mle2(y ~ dnorm(EVF(A,k,t=t,T_i),sd=stdv),#
        data = d,
       optimizer = "optimx",
        start = startlist4f,  
        parameters=list(A~1,k~1,T_i~treat),method="bobyqa",
      skip.hessian=FALSE))

ICtab(mod_gomp_1,mod_gomp_2,mod_EVF_1,mod_EVF_2,mod_gomp_2a,mod_gomp_2b,mod_gomp_2c,mod_gomp_2d,mod_gomp_2e,mod_gomp_2f,mod_EVF_2a,mod_EVF_2b,mod_EVF_2c,mod_EVF_2d,mod_EVF_2e,mod_EVF_2f,type="AICc",delta=T,nobs=nrow(d),weights=TRUE)
```



```{r}
d$pred <- predict(mod_gomp_2f)
mod_gomp_2f
ggplot(d, aes(x = t, y = y, color=treat)) +
  geom_point()+
  geom_function(fun = gompertz, args = list(A=coef(mod_gomp_2f)[1],k=coef(mod_gomp_2f)[2],T_i=coef(mod_gomp_2f)[3]),color="red")+
  geom_function(fun = gompertz, args = list(A=coef(mod_gomp_2f)[1],k=coef(mod_gomp_2f)[2],T_i=coef(mod_gomp_2f)[3]+coef(mod_gomp_2f)[4]),color="green")+
  geom_function(fun = gompertz, args = list(A=coef(mod_gomp_2f)[1],k=coef(mod_gomp_2f)[2],T_i=coef(mod_gomp_2f)[3]+coef(mod_gomp_2f)[5]),color="blue")
```
Zoom in on first 75 days
```{r}
ggplot(d, aes(x = t, y = y, color=treat)) +
  geom_point()+
  theme_classic() +
  scale_color_manual(values = c("olivedrab3", "cyan4", "purple4")) +
  geom_function(fun = gompertz, args = list(A=coef(mod_gomp_2f)[1],k=coef(mod_gomp_2f)[2],T_i=coef(mod_gomp_2f)[3]),color="olivedrab3")+
  geom_function(fun = gompertz, args = list(A=coef(mod_gomp_2f)[1],k=coef(mod_gomp_2f)[2],T_i=coef(mod_gomp_2f)[3]+coef(mod_gomp_2f)[4]),color="cyan4")+
  geom_function(fun = gompertz, args = list(A=coef(mod_gomp_2f)[1],k=coef(mod_gomp_2f)[2],T_i=coef(mod_gomp_2f)[3]+coef(mod_gomp_2f)[5]),color="purple4")+xlim(0,75)
```




```{r}
chickfullbci <- read.csv("~/Desktop/ECU/Research/Analysis/Weight & BCI/FullChickBCI.csv",na = c(""))

ggplot(chickfullbci, aes(x = day, y = bci, color=treat)) +
  geom_point()+
  scale_color_manual(values = c("olivedrab3", "cyan4", "purple4"))


require(tidyverse)

new.chickbci <- filter(chickfullbci) %>%
  mutate(age_group=cut(day, breaks = c(0,150,250),
             labels = c("Young","Old"), right = TRUE))

new.chickbci_young <- new.chickbci %>% 
  filter(age_group=="Young")

new.chickbci_old <- new.chickbci %>% 
  filter(age_group=="Old")


oldbcigraph <- ggplot(new.chickbci_old, aes(x = treat, y = bci, color=as.factor(treat))) +
  theme_classic() +
  scale_color_manual(values = c("olivedrab3", "cyan4", "purple4")) +
  geom_boxplot() +
  geom_point()

oldbcigraph



```
```{r}
chickfull <- read.csv("FullChickMeasurements.csv",na = c(""))

ggplot(chickfull, aes(x = day, y = weight, color=treat)) +
  geom_point()+geom_smooth()


new.chick <- filter(chickfull) %>%
  mutate(age_group=cut(day, breaks = c(0,150,250),
             labels = c("Young","Old"), right = TRUE))

new.chick_young <- new.chick %>% 
  filter(age_group=="Young")

new.chick_old <- new.chick %>% 
  filter(age_group=="Old")


oldgraph <- ggplot(new.chick_old, aes(x = treat, y = weight, color=as.factor(treat))) +
  theme_classic() +
  scale_color_manual(values = c("olivedrab3", "cyan4", "purple4")) +
  geom_boxplot() +
  geom_point()

oldgraph
```

